---
title: "ComputationalThinking1"
author: "Paige G and Leah W"
format: pdf
editor: visual
---

# Libraries

```{r}
library(tidyverse)
library(here)
```

## Question 1.1

```{r}
# Create a new function called add_together
# x and y will be the two arguments to the function
add_together  <- function(x, y){
  
  # Add x and y together, store as the object "output"
  output <- x + y
  
  # Print out whatever is stored in "output"
  return(output)
  
} 

add_together(x=3, y=5)
```

## Question 1.2

```{r eval= FALSE}
add_together(x=3, y="five")
```

```         
Error in x + y : non-numeric argument to binary operator
```

The error code is telling us that it's a numeric structure value, but it got a string instead.

## Question 1.3

```{r}
math_time <- function(x,y,z){
  output<- ((x-y)^2)/z
  return(output)
}

math_time(x=5,y=2,z=9)
```

## Question 1.4

```{r}
bison <- c(1000, 800, 1200, 1400)

deviation <- function(weights){
  mean<- mean(weights)
  sd <- weights-mean
  return(sd)
}

deviation(bison)
```

# Iteration

## Across function

```{r}
head(iris)
```

## Q 2.1

```{r}
help(iris)
```

The measurements are in centimeters.

## Q2.2

```{r}
iris %>%
  group_by(Species) %>% 
  summarize(across(.cols = everything(), 
                   .fns = median))
```

## Question 2.3

```{r}
cereal <- read.csv(here("data/cereal.csv"))
colnames(cereal)

cereal %>% 
  group_by(mfr) %>% 
  summarize(across(.cols = where(is.numeric),
                   .fns = mean))
```

## Q2.4

```{r}
for (i in 1:10) {
  print(sqrt(i))
}
```

## Q2.5a

```{r}
N0 = 300  #initial population size

years = 50  #number of years into the future

N = vector(length = years)  # create an empty vector to store pop. sizes

N[1] = N0  #initial population size should be the first N

lambda = 0.95  #growth rate

for (t in 2:50) {
  N[t] = N[t - 1] * lambda # Apply the equation
}

# store data
popn_data2 <- tibble(years = 1:years, # Make the years column = 1, 2, 3, ..., 20
                    popn = N) 
```

## Q2.5b

```{r}
popn_data2 %>% 
  ggplot(aes(x=years, y=popn))+
  geom_point()
```

## Q 2.6

Paige: Summarize/across makes the most sense, but probably has narrower applications than a for loop does. Having a strong understanding of a for loop has a lot of potential uses!

Leah: For loops make strong intuitive sense!

## Q 2.7

```{r}
# Store a vector of unique species names from the Species column of Iris
spp_names <- unique(iris$Species)

# Repeat the for loop for the number of unique species names. I.e., 3 species
# times
for (i in 1:length(spp_names)) {
  
  filt_data <- iris %>% 
    #  for each loop, create a filter where only one of the species names present.
    #creates 3 filt_data tibbles, one for each species
    filter(Species == spp_names[i])
  
  # for each loop, plot just that filtered data
  plot <- filt_data %>% 
    # map aesthetics to x and y axis using ggplot
    ggplot(aes(x = Petal.Length,
               y = Petal.Width)) +
    # scatter plot
    geom_point() +
    # with a line through it
    geom_smooth(method = "lm") +
    # set the x and y axis limits
    lims(x = c(0,8),
         y = c(0,3)) +
    # name the plot the species name that is specified for loop
    ggtitle(paste("Species:", spp_names[i]))
  
  # actually print/show the plot(s)
  print(plot)
}

```
